### 2015-06-05 LOI/inc2-old.tex
Major overhaul based on the Carlos, Matthew, and Jay conversation

{\bf Title of Pre-application:} Hierarchal Extreme Scale Knowledge Management \par
{\bf Principal Investigator:} Scott Klasky, Group Leader, ORNL, 854-241-9980, klasky@ornl.gov \par
{\bf Funding Opportunity Announcement Number: DE-FOA-0001338} \par
{\bf List of all co-PIs and Key/Senior Personnel} \par
Hasan Abbasi, Oak Ridge National Laboratory \par
Mark Ainsworth, Oak Ridge National Laboratory JFA, Brown University \par
Matthew Curry, Sandia National Laboratory\par
Qing Liu, Oak Ridge National Laboratory \par
Jay Lofstead, Sandia National Laboratory \par
Kimmy Mu, Oak Ridge National Laboratory \par
Carlos Malzahn, U. Cal. Santa Cruz \par
Manish Parashar, Rutgers University, Oak Ridge National Laboratory \par
Sudharshan Vazhkudai, Oak Ridge National Laboratory \par
Lee Ward, Sandia National Laboratory \par
\vskip 0.2 in

\underline{\textbf{Objectives:}} Exascale scientific discovery will
%introduce more complex hardware, and many simulations will 
be severely bottlenecked without sufficient new 
 research into managing and storing the large amounts of 
data that will be produced during the simulation, and analyzed for months after the simulations.
%Our goal in this project is to investigate a unique approach to addressing this data management 
%and storage challenge for multi-tier exascale storage architectures, thereby expediting insights into mission critical
%scientific processes. 
Our goal in this project is to address this data management and storage challenge. Specifically we plan
 to research and develop a unique application-aware hierarchy data organization that can effectively map 
 to  multi-tier exascale storage architectures, and can facilitate how data can be efficiently written to and 
 read from a complex heterogeneous storage hierarchy  
 thereby expediting insights into mission critical scientific processes. 
% the multi-tier challenges that are faced by scientists in creating and
% managing and storing their data  to expedite insights xsinto
% mission critical scientific processes in exascale compuxsting.
%
%Specifically, we propose a research program that aims to explore a hierarchical
%organization and storage infrastructure that will facilitate how data can
%be written to and read from a complex heterogenous storage hierarchy
%efficiently.  
%
Our approach is to explicitly and implicitly capture user/application intentions 
and use this information, along with direct feedback from the storage and I/O 
system, to dynamically structure, organize, process and map the data so that 
it can most effectively support complex I/O patterns. Furthermore, we will do 
this with an awareness of data usage semantics and with an understanding 
of current and future hardware and software technologies. 
%re-organize and possibly reduce the total amount of information 
%Since data analysis and visualization will be performed in both an
%in situ and post processing environment, we will continue to use our I/O and
%storage abstractions to help enable applications to take full advantage of
%the hardware and software environments. 
For example, as data volumes are quickly outgrowing available bandwidths, 
compression is becoming essential. Rather than focusing solely on lossless 
compression, we will incorporate lossy compression with the ability to annotate 
data allowing optimized algorithm selection based on the numerical methods 
employed in the source simulation. We plan to explore this tradeoff of storage 
capacity and bandwidth for computation to determine how to offer the user 
options that can meet their needs with desirable performance characteristics.

In order to explore the many research challenges in this domain, we will build 
upon the success of our middleware system, ADIOS, our multi-tier storage 
system, Sirocco, and our distributed storage system, Ceph.  We will also leverage 
data staging and data movement technologies, DataSpaces and the Common 
Communication Interface (CCI), as part of a software stack. 
%used successfully on current leadership class machines. 
%Our experience with this stack will guide us in understanding the new directions for
%research in storage systems and I/O for exascale and beyond systems.

@@

\underline{\textbf{Key Technical Approach:}}
%We are fundamentally trying to address the fundmantal questions : 1)  How can we 
%place and mangage massive scienfici data across all of the tiers of the storage and memory
%hiearchy? 2) What are the proper semantics needed in order to help bring knowledge from the applications
%to the middleware layer. 3) What information from the storage layer can be exposed to the middleweare such
%that the placement of information can be agreed upon from what the application requires and what the
%system resources are available. 
Our overall technical approach is based on an application-aware runtime realization of tradeoffs in 
data-representation, data placement and data access. For example, we will allow users to ``plug-in'' 
techniques to classify data, not as bytes but as motifs, allowing us to understand relationships 
between objects (based on data models), the relationships of data and variables, as well as data
access and transport patterns.  This will allow us 
to adaptively map data from a variable onto the various storage tiers. Rather than focus on simple data
compression, we will incorporate a multitude of techniques to reduce data on the faster storage tiers, and keep the less
reduced information in the lower storage tiers.

>>>>

{\bf Title of Pre-application:} Hierarchal Extreme Scale Knowledge Management \par
{\bf Principal Investigator:} Scott Klasky, Group Leader, ORNL, 854-241-9980, klasky@ornl.gov \par
{\bf Funding Opportunity Announcement Number: DE-FOA-0001338} \par
{\bf List of all co-PIs and Key/Senior Personnel} \par
Hasan Abbasi, Oak Ridge National Laboratory \par
Mark Ainsworth, Oak Ridge National Laboratory JFA, Brown University \par
Matthew Curry, Sandia National Laboratory\par
Qing Gary Liu, Oak Ridge National Laboratory \par
Jay Lofstead, Sandia National Laboratory \par
Kimmy Mu, Oak Ridge National Laboratory \par
Carlos Malzahn, U. Cal. Santa Cruz \par
Manish Parashar, Rutgers University, Oak Ridge National Laboratory \par
Sudharshan Vazhkudai, Oak Ridge National Laboratory \par
Lee Ward, Sandia National Laboratory \par
\vskip 0.2 in

\underline{\textbf{Objectives:}} Exascale scientific discovery will
%introduce more complex hardware, and many simulations will 
be severely bottlenecked without sufficient new 
 research into managing and storing the large amounts of 
data that will be produced during the simulation, and analyzed for months afterwards.
%Our goal in this project is to investigate a unique approach to addressing this data management 
%and storage challenge for multi-tier exascale storage architectures, thereby expediting insights into mission critical
%scientific processes. 
Our goal in this project is to address the associated I/O and storage challenges on the 
future storage landscape, and expedite insights into mission critical
scientific processes. To this end, we plan
 to research and develop a unique application-aware data management scheme that
can (1) effectively map data to multi-tier exascale storage architectures, 
(2) facilitate how data can be efficiently written to and 
 read from a complex heterogeneous storage hierarchy, and (3) reduce and re-generate
data on-the-fly without compromising the quality of science. The project
will assemble a team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL, UCSC)
and storage (UCSC), and connect and coordinate
these key storage components in a seamless fashion. 
% the multi-tier challenges that are faced by scientists in creating and
% managing and storing their data  to expedite insights xsinto
% mission critical scientific processes in exascale compuxsting.
%
%Specifically, we propose a research program that aims to explore a hierarchical
%organization and storage infrastructure that will facilitate how data can
%be written to and read from a complex heterogenous storage hierarchy
%efficiently.  
%
Our approach is to explicitly and implicitly capture user/application intentions 
and use this information, along with direct feedback from the I/O and storage
system, to dynamically structure, organize, process and map the data so that 
it can most effectively support complex I/O patterns. This will be done
with an awareness of data usage semantics and an understanding 
of current and future hardware and software technologies. 
%re-organize and possibly reduce the total amount of information 
%Since data analysis and visualization will be performed in both an
%in situ and post processing environment, we will continue to use our I/O and
%storage abstractions to help enable applications to take full advantage of
%the hardware and software environments. 
On the other hand, data reduction is becoming increasingly important, as a result of 
the disparirty between data volume/frequency and storage bandwdith. 
Rather than focusing solely on lossless 
compression, we will incorporate lossy compression with the ability to annotate 
data, allowing for optimized algorithm selection based on the numerical accuracy
required by the simulation and analysis. We plan to explore this tradeoff of storage 
capacity and bandwidth for computation to determine how to offer the user 
options that can meet their needs with desirable performance characteristics.

%In order to explore the many research challenges in this domain, we will build 
%upon the success of our middleware system, ADIOS, our multi-tier storage 
%system, Sirocco, and our distributed storage system, Ceph.  We will also leverage 
%data staging and data movement technologies, DataSpaces and the Common 
%Communication Interface (CCI), as part of the I/O stack. 
%%used successfully on current leadership class machines. 
%Our experience with this stack will guide us in understanding the new directions for
%research in storage systems and I/O for exascale and beyond systems.

@@

\underline{\textbf{Key Technical Approach:}}
%We are fundamentally trying to address the fundmantal questions : 1)  How can we 
%place and mangage massive scienfici data across all of the tiers of the storage and memory
%hiearchy? 2) What are the proper semantics needed in order to help bring knowledge from the applications
%to the middleware layer. 3) What information from the storage layer can be exposed to the middleweare such
%that the placement of information can be agreed upon from what the application requires and what the
%system resources are available. 
Our overall technical approach is based on an application-aware runtime realization of tradeoffs in 
data representation, data placement and data access. We will allow users to ``plug-in'' 
their knowledge about the data, not as bytes but as motifs, allowing the I/O and storage system
to understand user intentions and the relationships between data objects, as well as data
access and transport patterns. This will facilitate
efficient mapping of data from user space, such as a multi-dimensional array, onto various storage tiers. 
Rather than focus on simple data
compression, we will incorporate a multitude of techniques to reduce data on the faster yet smaller
storage tiers, and keep the less
reduced information in the lower storage tiers.

### 2015-06-05 LOI/inc2.tex
Major overhaul based on the Carlos, Matthew, and Jay conversation

{\bf Title of Pre-application:} Hierarchal Extreme Scale Knowledge Management \par
{\bf Principal Investigator:} Scott Klasky, Group Leader, ORNL, 854-241-9980, klasky@ornl.gov \par
{\bf Funding Opportunity Announcement Number: DE-FOA-0001338} \par
{\bf List of all co-PIs and Key/Senior Personnel} \par
Hasan Abbasi, Oak Ridge National Laboratory \par
Mark Ainsworth, Oak Ridge National Laboratory JFA, Brown University \par
Matthew Curry, Sandia National Laboratory\par
Qing Liu, Oak Ridge National Laboratory \par
Jay Lofstead, Sandia National Laboratory \par
Kimmy Mu, Oak Ridge National Laboratory \par
Carlos Malzahn, U. Cal. Santa Cruz \par
Manish Parashar, Rutgers University, Oak Ridge National Laboratory \par
Sudharshan Vazhkudai, Oak Ridge National Laboratory \par
Lee Ward, Sandia National Laboratory \par
\vskip 0.2 in

\underline{\textbf{Objectives:}} Exascale scientific discovery will
%introduce more complex hardware, and many simulations will 
be severely bottlenecked without sufficient new 
 research into managing and storing the large amounts of 
data that will be produced during the simulation, and analyzed for months after the simulations.
%Our goal in this project is to investigate a unique approach to addressing this data management 
%and storage challenge for multi-tier exascale storage architectures, thereby expediting insights into mission critical
%scientific processes. 
Our goal in this project is to address this data management and storage challenge. Specifically we plan
 to research and develop a unique application-aware hierarchy data organization that can effectively map 
 to  multi-tier exascale storage architectures, and can facilitate how data can be efficiently written to and 
 read from a complex heterogeneous storage hierarchy  
 thereby expediting insights into mission critical scientific processes. 
% the multi-tier challenges that are faced by scientists in creating and
% managing and storing their data  to expedite insights xsinto
% mission critical scientific processes in exascale compuxsting.
%
%Specifically, we propose a research program that aims to explore a hierarchical
%organization and storage infrastructure that will facilitate how data can
%be written to and read from a complex heterogenous storage hierarchy
%efficiently.  
%
Our approach is to explicitly and implicitly capture user/application intentions 
and use this information, along with direct feedback from the storage and I/O 
system, to dynamically structure, organize, process and map the data so that 
it can most effectively support complex I/O patterns. Furthermore, we will do 
this with an awareness of data usage semantics and with an understanding 
of current and future hardware and software technologies. 
%re-organize and possibly reduce the total amount of information 
%Since data analysis and visualization will be performed in both an
%in situ and post processing environment, we will continue to use our I/O and
%storage abstractions to help enable applications to take full advantage of
%the hardware and software environments. 
For example, as data volumes are quickly outgrowing available bandwidths, 
compression is becoming essential. Rather than focusing solely on lossless 
compression, we will incorporate lossy compression with the ability to annotate 
data allowing optimized algorithm selection based on the numerical methods 
employed in the source simulation. We plan to explore this tradeoff of storage 
capacity and bandwidth for computation to determine how to offer the user 
options that can meet their needs with desirable performance characteristics.

In order to explore the many research challenges in this domain, we will build 
upon the success of our middleware system, ADIOS, our multi-tier storage 
system, Sirocco, and our distributed storage system, Ceph.  We will also leverage 
data staging and data movement technologies, DataSpaces and the Common 
Communication Interface (CCI), as part of a software stack. 
%used successfully on current leadership class machines. 
%Our experience with this stack will guide us in understanding the new directions for
%research in storage systems and I/O for exascale and beyond systems.

% We will also take advantage of our staging system, Data Spaces, and the
% network abstraction layer, the Common communication Interface (CCI),
% and build on all of
% our sucessful software used in petascale simulations to understand the new directions which need to be understood to fully develop
% the next generation I/O and Storage System for exascale and beyond systems.

%

%
Our objective here is to reduce the time to knowledge, an end-to-end metric
relevant to the scientific discovery process. Beyond the traditional high
volume I/O pattern of checkpoint/restart, we will address the challenges
posed by other essential data access patterns in the knowledge gathering
process. Through a deeper insight into the scientific process we will encode
and utilize accuracy and errors as optimization parameters. 
%
% We will explore access to
% large data sets as part of this process, w

% which means that we will
% also support additional access modes, beyond the traditional
% Checkpoint/Restart methods, and understand that all data has errors
% associated with it and we will ensure that these become encoded to allow for
% additional optimizations. 
For example, scientific simulations contain approximations, as do measurements 
from observations and experiments, and depending on the goal, these approximations 
are acceptable. We can leverage this observation to optimize the presentation of data
to the user and to implement various tradeoffs. For example, users can ask for information 
within a given accuracy bound, allowing us to offer a mode for re-computation vs. data storage
and reterival.

%
{\bf: please list other things from UCSC, OLCF}.
%

\underline{\textbf{Key Technical Approach:}}
%We are fundamentally trying to address the fundmantal questions : 1)  How can we 
%place and mangage massive scienfici data across all of the tiers of the storage and memory
%hiearchy? 2) What are the proper semantics needed in order to help bring knowledge from the applications
%to the middleware layer. 3) What information from the storage layer can be exposed to the middleweare such
%that the placement of information can be agreed upon from what the application requires and what the
%system resources are available. 
Our overall technical approach is based on an application-aware runtime realization of tradeoffs in 
data-representation, data placement and data access. For example, we will allow users to ``plug-in'' 
techniques to classify data, not as bytes but as motifs, allowing us to understand relationships 
between objects (based on data models), the relationships of data and variables, as well as data
access and transport patterns.  This will allow us 
to adaptively map data from a variable onto the various storage tiers. Rather than focus on simple data
compression, we will incorporate a multitude of techniques to reduce data on the faster storage tiers, and keep the less
reduced information in the lower storage tiers.

We will support additional data access modes as well. First, within a given
accuracy bounds, offer a mode for re-computation from data stored in a ``fast''
tier to reduce data latencies. Second, exploratory analysis operations
frequently entail an overall data set view followed by targeted data
exploration based on identified features. We will offer support for a
configurable data access mode to support these sorts of analysis accesses. An
``overview'' access mode that gives a quick, approximate within error bounds,
data view that can guide feature selection offering rapid coarse-grained data
exploration without requiring loading the complete, detailed data set from
storage. Based on the granularity requested, the accuracy and size of the data
returned can be adjusted. At the most extreme setting, the original data can
be retrieved at the time cost of moving the potentially huge data quantity.

Third, to ensure available storage for subsequent operations, we will offer
automatic data migration based on user annotations for required data lifetimes
using monitoring and learning techniques. Unlike existing approaches, this will
be tempered both by the user annotations and through learned access patterns.
While past access patterns may not indicate future access because the
simulation run purpose may have changed, we are focused on scalability where
runs are subsequently larger as the simulation prepares for a capability run.
By learning from the output and access patterns during this run sequence, we
can accurate decide how to place and organize data for the critical capability
runs. Fourth, we anticipate storing multiple data copies, each compressed in
different ways according to the underlying media, some of these copies will
disappear based on storage pressures, but data persistence will be maintained
according to user specifications. Assuming a relatively low latency cache layer
before a tape system, we can offer exploratory data access reserving pulling
data from tape to just the data required. This will save scientists time and
make data stored on tape usable without long delays.

@@

%We will identify,  through concrete application evaluation, the requirements for highly usable and scalable middleware and storage layers

>>>>

{\bf Title of Pre-application:} Hierarchal Extreme Scale Knowledge Management \par
{\bf Principal Investigator:} Scott Klasky, Group Leader, ORNL, 854-241-9980, klasky@ornl.gov \par
{\bf Funding Opportunity Announcement Number: DE-FOA-0001338} \par
{\bf List of all co-PIs and Key/Senior Personnel} \par
Hasan Abbasi, Oak Ridge National Laboratory \par
Mark Ainsworth, Oak Ridge National Laboratory JFA, Brown University \par
Matthew Curry, Sandia National Laboratory\par
Qing Gary Liu, Oak Ridge National Laboratory \par
Jay Lofstead, Sandia National Laboratory \par
Kimmy Mu, Oak Ridge National Laboratory \par
Carlos Malzahn, U. Cal. Santa Cruz \par
Manish Parashar, Rutgers University, Oak Ridge National Laboratory \par
Sudharshan Vazhkudai, Oak Ridge National Laboratory \par
Lee Ward, Sandia National Laboratory \par
\vskip 0.2 in

\underline{\textbf{Objectives:}}
Exascale scientific discovery will be severely bottlenecked without sufficient
new research into managing and storing the large amounts of data that will be
produced during the simulation, and analyzed for months afterwards.  Our goal
in this project is to address the associated I/O and storage challenges on the
future storage landscape, and expedite insights into mission critical
scientific processes. To that end, we will build on the capabilities offered
by the Sirocco peer-to-peer file system under development at Sandia and the
object storage and annotation expertise of UC Santa Cruz to develop a unique
application aware data management scheme. It will build on the existing
multi-tier storage aware approach offered by Sirroco to incorporate new
functionality and extended object metadata and APIs for 1) matching available
storage bandwidth across the storage hierarchy with data importance employing
application aware data compression (both lossy with specified error bounds and
lossless), 2) a selectable performance/quality tradeoff when reading data, and
3) incorporate reader prioritization for data annotation, placement, and data
quality when writing data.  The major challenge in this project is metadata
management connecting different object quality levels and connecting the user
with the best possible outcomes given the current system state. The new user
APIs required to interact with this rich metadata system will drive effective
use of the entire middleware and storage stack. The project is compriesed of a
team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL,
UCSC) and storage (UCSC), and connect and coordinate these key storage
components in a seamless fashion.

Our approach can most easily be understood by considering a road navigation
program. Similar to the navigation program, we will offer a coarse grained,
fast overview of a data set with progressively more detail in areas of interest
noted by data features and less data in areas without. These progressively
more detailed data views require more storage space and time to retrieve. By
incorporating middleware informed of both how to identify relevant data
features and the intended use locations, it will be able to selectively store
data at different quality levels, with bounded errors in the case of lossy
compression, and allow a user to select what data to retrieve by specifying
both an acceptable retrieval timeframe and quality required. If the request
cannot be fulfilled, the storage system can negotiate with the user providing
information about the quality and timeframes available for a subsequent
request.

Our objective here is to reduce the time to knowledge, an end-to-end metric
relevant to the scientific discovery process. Beyond the traditional high
volume I/O pattern of checkpoint/restart, we will address the challenges posed
by other essential data access patterns in the knowledge gathering process.
Through a deeper insight into the scientific process we will encode and utilize
accuracy and errors as optimization parameters.

For example, scientific simulations contain approximations, as do measurements
from observations and experiments, and depending on the goal, these
approximations are acceptable. We can leverage this observation to optimize the
presentation of data to the user and to implement various tradeoffs. For
example, users can ask for information within a given accuracy bound, allowing
us to offer a mode for re-computation vs. data storage and reterival.

%
{\bf: please list other things from OLCF}.
%

\underline{\textbf{Key Technical Approach:}}
Our overall technical approach is based on an application-aware runtime
realization of tradeoffs in data representation, data placement and data
access. We will allow users to ``plug-in'' their knowledge about the data, not
as bytes but as motifs, allowing the I/O and storage system to understand user
intentions and the relationships between data objects, as well as data access
and transport patterns. This will facilitate efficient mapping of data from
user space, such as a multi-dimensional array, onto various storage tiers.
Rather than focus on simple data compression, we will incorporate a multitude
of techniques to reduce data on the faster yet smaller storage tiers, and keep
the less reduced information in the lower storage tiers.

We will support additional data access modes as well. \textbf{First}, within a
given accuracy bounds, offer a mode for re-computation from data stored in a
``fast'' tier to reduce data latencies. \textbf{Second}, exploratory analysis
operations frequently entail an overall data set view followed by targeted data
exploration based on identified features. We will offer support for a
configurable data access mode to support these sorts of analysis accesses. An
``overview'' access mode that gives a quick, approximate within error bounds,
data view that can guide feature selection offering rapid coarse-grained data
exploration without requiring loading the complete, detailed data set from
storage. Based on the granularity requested, the accuracy and size of the data
returned can be adjusted. At the most extreme setting, the original data can
be retrieved at the time cost of moving the potentially huge data quantity.
\textbf{Third}, to ensure available storage for subsequent operations, we will
offer automatic data migration based on user annotations for required data
lifetimes using monitoring and learning techniques. Unlike existing approaches,
this will be tempered both by the user annotations and through learned access
patterns.  While past access patterns may not indicate future access because
the simulation run purpose may have changed, we are focused on scalability
where runs are subsequently larger as the simulation prepares for a capability
run.  By learning from the output and access patterns during this run sequence,
we can accurate decide how to place and organize data for the critical
capability runs. Fourth, we anticipate storing multiple data copies, each
compressed in different ways according to the underlying media, some of these
copies will disappear based on storage pressures, but data persistence will be
maintained according to user specifications. Assuming a relatively low latency
cache layer before a tape system, we can offer exploratory data access
reserving pulling data from tape to just the data required. This will save
scientists time and make data stored on tape usable without long delays.

@@
